---
title: "CAMFAM: An Analysis of Suicide Rates"
date: May 1st, 2019
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(dslabs)
library(nnet)
library(knitr)
library(broom)
library(skimr)
library(modelr)
library(leaps)
library(plyr)
library(dplyr)
library(rms)
```

```{r load-data, include = FALSE}
suicide <- read_csv("/cloud/project/data/master.csv")
```
## Introduction

With suicide being one of the leading causes of death for teens in the world, it is important to not only see how Northern America (Canada, Greenland, and the United States) stack up with the rest of the world, but also the key trends of suicide rates among different groups and what factors shed light onto these different rates.[1] We want to explore how economic status, along with variables such as age, sex, and human development index, affects suicide rates all across the world. Our hypothesis is that generally, in poorer countries we predict that suicide rates will be lower.

Our response variable will be suicides/100k pop, which is the number of suicides per 100,000 people in a certain country and year and is stored as a numeric in our dataset. Our predictor variables will be age, sex, HDI, gdp_for_year, gdp_per_capita, generation, population, region and continent. Age is the age an individual was when they passed, sex is the gender of that individual, country is the country they are from, year is the year they passed, HDI for year is the human development index for a given country and year, gdp_for_year is the GDP for a given country and year, gdp_per_capita is the GDP per capita for a given country and year, and generation is the generation that an individual belongs to. There generations are Silent (born 1928-1945), Generation X (born 1965-1980), Millennials (born 1981-1996), and Generation Z (born 1997-2015). We wish to understand how the number of suicides per 100,000 people in a certain country and year changes as population, GDP, GDP per capita, and HDI increase or decreases, meaning we want to understand the population coefficients for gdp_for_year, gdp_per_capita, and HDI for year. Additionally, we want to understand whether age, sex, generation, and location have an effect on the number of suicides per 100,000 people, meaning we also want to understand the population coefficients for these variables. We will not include country and suicide/no in our analysis because having 188 different levels is unrealistic for the former, and the latter is manifested in the response variable. 

### EDA
```{r, echo = FALSE}
suicideNew <- suicide%>% filter(year == 2010)
suicideNew <- suicideNew %>% mutate(age = as.factor(age))
suicideNew <- suicideNew %>% mutate(country = as.factor(country))
suicideNew <- suicideNew %>% mutate(generation = as.factor(generation))
suicideNew <- suicideNew %>% mutate(sex = as.factor(sex))
gapminder2010 <- gapminder %>% filter(year == 2010)
gapminder2010 <- gapminder2010 %>% select(country, continent, region)
gapminder2010 <- data.frame(lapply(gapminder2010, as.character), stringsAsFactors=FALSE)
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="Kyrgyz Republic", "Kyrgyzstan")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="Slovak Republic", "Slovakia")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="South Korea", "Republic of Korea")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="Russia", "Russian Federation")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="St. Lucia", "Saint Lucia")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="St. Vincent and the Grenadines", "Saint Vincent and Grenadines")
gapminder2010 <- data.frame(lapply(gapminder2010, as.character), stringsAsFactors=TRUE)
suicideFinal <- merge(suicideNew, gapminder2010, by = "country", all.x = TRUE)
names(suicideFinal)[names(suicideFinal) == 'HDI for year'] <- 'HDI'
group.means<-ddply(suicideFinal,c("region"),summarise,mean=mean(HDI, na.rm = TRUE))
for(i in 1:nrow(suicideFinal)){
  if (suicideFinal$country[i] == "Republic of Korea"){
    suicideFinal$HDI[i] = .884
  }
  if (suicideFinal$country[i] == "Russian Federation"){
    suicideFinal$HDI[i] = .780
  }
  if (suicideFinal$country[i] == "Aruba" | suicideFinal$country[i] == "Puerto Rico"){
    suicideFinal$HDI[i] = .751
  }
}
suicideFinal <- suicideFinal %>% mutate (`suicides_no` = `suicides_no`+1)
suicideFinal <- suicideFinal %>% mutate (`suicides/100k pop` = `suicides_no`/(population/100000))
```

Given the longitudinal structure of our data and the need for making a multilevel model, we are going to use the data collected in 2010. This year has a majority of the countries in the origianl data set and also is one of the years HDI values were collected. Since region and continent were not in this data set orginally, we used the gapminder data set as a cold deck and merged those values into the suicide data set.

Since South Korea and Russia did not have HDI values, we found them online and imputed those (they had values of .884 and .780 respectively) as well. However, Aruba and Puerto Rico's HDI's were not available, so we took the average of the HDI in the Caribbean and used a mean imputation for those values. 

For the response variable, since there are a few negative infinity values, we added one to each number of suicides, divided by population (to turn into suicided/100k) and then log transformed. We log transformed because of the original skewed histogram (Figure 1).

```{r message = FALSE, warning = FALSE, echo = FALSE}
suicideFinal <- suicideFinal %>% mutate (`suicides/100k pop` = log(`suicides/100k pop`))
```

```{r, echo = FALSE}
suicideBar<-ggplot(data=suicideFinal, mapping=aes(x=`suicides/100k pop`)) +
  geom_histogram() +
  labs(title="Distribution of Suicides", x="Log Number of Suicides per 100,000 People")

continentBar <- ggplot(data=suicideFinal, mapping=aes(x=`continent`)) + 
  geom_bar() +
  labs(title="Bar Plot of Continents", x="Continent") + 
  theme(axis.text.x = element_text(angle=90))

regionBar <- ggplot(data=suicideFinal, mapping=aes(x=`region`)) + 
  geom_bar() +
  labs(title="Bar Plot of Regions", x="Region") + coord_flip() +  theme(axis.text.x = element_text(angle=90))


continentBox <- ggplot(data=suicideFinal, mapping=aes(x=`continent`, y= `suicides/100k pop`)) + 
  geom_boxplot() +
  labs(title="Continent vs. Suicides", x="Continent", y="Suicides") +
 theme(axis.text.x = element_text(angle=90))

require(cowplot)
plot_grid(suicideBar, continentBar, regionBar, continentBox)
```

The mean number of suicides per 100,000 people is 11.22, while the maximum number of suicides per 100,000 people in this dataset is 182.32. The distribution of this reponse variable is normal. We see that most of our countries are from the Amerias and Europe and are in the regions of Western Asia, Southern Europe, South America, Northern Europe, and the Caribbean. In general, each continent has a similar average number of suicides, with Europe having much more outliers with fewer suicides than the others.

```{r, echo = FALSE}
ageBox <- ggplot(data=suicideFinal, mapping=aes(x=`age`, y= `suicides/100k pop`)) + 
  geom_boxplot() +
  labs(title="Age vs. Suicides", x="Age", y="Suicides") +
 theme(axis.text.x = element_text(angle=90))

regionBox <- ggplot(data=suicideFinal, mapping=aes(x=`region`, y= `suicides/100k pop`)) + 
  geom_boxplot() +
  labs(title="Region vs. Suicides", x="Region", y="Suicides") + coord_flip()

sexBox <- ggplot(data=suicideFinal, mapping=aes(x=sex, y=`suicides/100k pop` ))+
  geom_boxplot()+
  labs(title = "Boxplot of Gender & Suicide", x = "gender", y = "suicide")

generationBox <- ggplot(data=suicideFinal, mapping=aes(x=generation, y=`suicides/100k pop` ))+
  geom_boxplot()+
  labs(title = "Boxplot of Generation & Suicide", x = "Generation", y = "Suicide") +
 theme(axis.text.x = element_text(angle=90))

require(cowplot)
plot_grid(ageBox, regionBox, generationBox, sexBox)
```

As age increases, suicide rate tends to increase in general. The 5-14 age group is far lower than the other age groups, which is expected. The average number of suicides/100k for each region is around 7.38. We see Southern Africa as a key outlier with far less suicides and Eastern Africa with far more suicides than the average. We notice that males have a higher suicide rate/100k people than females. However, there are many outliers in this data set so we must explore further.
We notice that Generation Z is significantly lower in terms of the average number of suicides/100k of 7.38.

Due to the skewed histograms of GDP, GDP per capita, and population, we log transformed them (Figures 2,3,4). We also mean center all the quantitative variables.

```{r, echo = FALSE, fig.height = 4}
suicideFinal <- suicideFinal %>% mutate(population = log(population), population = population - mean(population))
popHist <- ggplot(data=suicideFinal, mapping=aes(x=`population`)) + 
  geom_histogram() +
  labs(title="Distribution of Population", x="Population")

suicideFinal <- suicideFinal %>% mutate(HDI = HDI - mean(HDI))
hdiHist <- ggplot(data=suicideFinal, mapping=aes(x=`HDI`)) + 
  geom_histogram() +
  labs(title="Distribution of HDI", x="HDI")

suicideFinal<- suicideFinal %>% mutate (`gdp_for_year ($)` = log(`gdp_for_year ($)`))
suicideFinal <- suicideFinal %>% mutate(`gdp_for_year ($)` = `gdp_for_year ($)` - mean(`gdp_for_year ($)`))
gdpHist <- ggplot(data=suicideFinal, mapping=aes(x=(`gdp_for_year ($)`))) + geom_histogram() +
  labs(title="Distribution of GDPs of Countries", x="Value")

suicideFinal<- suicideFinal %>% mutate (`gdp_per_capita ($)` = log(`gdp_per_capita ($)`))
suicideFinal <- suicideFinal %>% mutate(`gdp_per_capita ($)` = `gdp_per_capita ($)` - mean(`gdp_per_capita ($)`))
gdpcapHist <- ggplot(data=suicideFinal, mapping=aes(x=(`gdp_per_capita ($)`))) +
  geom_histogram() +
  labs(title="Distribution of GDP Per Capita", x="Value")

require(cowplot)
plot_grid(gdpHist, gdpcapHist, popHist, hdiHist)
```

The HDI does not have a set distribution. Rather, it is quite sparse. The distribution of population is skewed left and has a unimodal distribution. We notice that the distribution of GDPs of countries has a near-bimodal distribution.The GDP Per Capita is skewed to the left and is bimodal.

```{r, echo = FALSE, fig.height = 4}
hdiDist <- ggplot(data=suicideFinal, mapping=aes(x=`HDI`, y=`suicides/100k pop`))+
  geom_point()+
  labs(title = "Plot of Suicides/100k vs HDI", x = "HDI", y = "Suicides/100k")

popDist <- ggplot(data=suicideFinal, mapping=aes(x=`population`, y=`suicides/100k pop`))+
  geom_point()+
  labs(title = "Plot of Suicides/100k vs Pop", x = "Population", y = "Suicides/100k")

gdpcapDist <- ggplot(data=suicideFinal, mapping=aes(x=`gdp_per_capita ($)`, y=`suicides/100k pop`))+
  geom_point()+
  labs(title = "Plot of Suicides/100k vs GDP Per Capita", x = "GDP Per Capita", y = "Suicides/100k")

gdpDist <- ggplot(data=suicideFinal, mapping=aes(x=`gdp_for_year ($)`, y=`suicides/100k pop`))+
  geom_point()+
  labs(title = "Plot of Suicides/100k vs GDP", x = "GDP", y = "Suicides/100k")

require(cowplot)
plot_grid(hdiDist, popDist, gdpDist, gdpcapDist)
```

We do not see correlations between HDI, GDP, and GDP per capita and suicides/100k. We notice that there could some sort of linear trend or threshold for population and suicides. Perhaps in those countries with extremely low populations the number of suicides/100k is generally higher.

```{r, fig.height = 4, echo = FALSE}
pairs(`suicides/100k pop` ~ `HDI` + `gdp_for_year ($)` + `gdp_per_capita ($)` + population, data = suicideFinal)
```

From the pairs plot, it looks as if HDI, and gdp_for_year do not have a clear linear relationship with suicides/100k pop. However, gdp_per_capita seems to be positively correlated with suicides/100k pop, meaning as gdp_per_capita increases, so does suicides/100k pop. Additionally, it looks as if HDI and gdp_per_capita seem to have a strong non-linear relationship, indicating that we should continue looking into this relationship and perhaps include an interaction term between these two variables. Similarly, HDI and gdp_for_year also seem to have a strong non-linear relationship, so we should include an interaction term between these two variables as well. We also see a strong evidence of multicollinearity between GDP for year and GDP per capita that we must address in the model.

We are very concerned about multicollinearity between all of these variables, so we will look into VIF.

We plan to do a multiple linear regression because suicides/100k pop is a quantitative variable (there are no levels to it, since it is a continuous variable).


## Regression Analysis

```{r, echo = FALSE}
model1 <- lm(`suicides/100k pop` ~ sex + age + population + HDI + (`gdp_for_year ($)`) + (`gdp_per_capita ($)`) + continent + region + generation, data = suicideFinal)
regfit_backward <- step(model1, direction = "backward")
```
After conducting a backward selection, we we found that sex, age population, HDI, gdp for year, and gdp per capita, and region are relevant predictors, whereas population, continent, and generation are not.

### Interesting Interactions
```{r}
regionPop <- ggplot(data=suicideFinal,mapping=aes(x=population,y=`suicides/100k pop`,color=as.factor(region))) +
geom_smooth(method="lm", se=FALSE) + labs(title="Population & Region Interaction", 
x="population", 
y="suicides",
color="region") + theme(legend.position = "none")

regionHdi <-  ggplot(data=suicideFinal,mapping=aes(x=HDI,y=`suicides/100k pop`,color=as.factor(region))) +
geom_smooth(method="lm", se=FALSE) + labs(title="HDI & Region Interaction", 
x="HDI", 
y="suicides",
color="region") + theme(legend.position = "none")

regionGDPCAP <- ggplot(data=suicideFinal,mapping=aes(x=`gdp_per_capita ($)`,y=`suicides/100k pop`,color=as.factor(region))) +
geom_smooth(method="lm", se=FALSE) + labs(title="GDP/Capita & Region Interaction", 
x="GDP / capita", 
y="suicides",
color="region") + theme(legend.position = "none")

regionGDP <- ggplot(data=suicideFinal,mapping=aes(x=`gdp_for_year ($)`,y=`suicides/100k pop`,color=as.factor(region))) +
geom_smooth(method="lm", se=FALSE) + labs(title="GDP & Region Interaction", 
x="GDP / year", 
y="suicides",
color="region") + theme(legend.position = "none")

legend_b <- get_legend(regionGDP + theme(legend.position="bottom"))

require(cowplot)
plots <- plot_grid(regionPop, regionHdi, regionGDPCAP, regionGDP)
plot_grid(plots, legend_b, ncol = 1, rel_heights = c(1, .1))
```


From all of these plots, we see interactions between each qualitative and quantitative variable tested. However, after conducting nested F tests, we see that there are interaction effects between all the above interactions except region and population. There were no interaction effects between sex and any quantitative variable.

```{r, echo = FALSE}
finalmodel <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + region*HDI + region*`gdp_for_year ($)` + region*`gdp_per_capita ($)` + age*HDI + age*population + age*`gdp_per_capita ($)` + age*`gdp_for_year ($)`, data = suicideFinal)
```

```{r, echo = FALSE,  warning = FALSE}
suicideFinal <- suicideFinal %>%
  mutate(predicted = predict.lm(finalmodel), resid = residuals(finalmodel))

resids<- ggplot(data=suicideFinal, mapping=aes(x=predicted,y=resid)) +
  geom_point() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. Predicted Values",
       x="Predicted Values", 
       y="Residuals")

qq<- ggplot(data=suicideFinal, mapping=aes(sample=resid)) +
  stat_qq() + 
  stat_qq_line()+
labs(title="Normal QQ Plot of Residuals")

distResids <- ggplot(data=suicideFinal, mapping=aes(x=resid)) +
  geom_histogram() +  
  labs(title="Distribution of the Residuals", x="Residuals")

plot_grid(resids, qq, distResids)
```

From the residual vs. predicted values, the residuals with lower predicted values seem to be more sparse and spread out than the others, which may be worth looking into. However, for the most part, the majority of the residuals are randomly distributed around the red line and do not exhibit any obvious nonlinear trends. Thus, we can conclude that constant variance is satisfied. According to the QQ-plot, the normality assumption seems to be satisfied because the majority of the points seem to align well with the expected QQ-plot line. In addition, the histogram seems to be approximately normal and follows a mostly smooth curve. Based on the description of the data, the independence assumption seems to be met because this data was not collected over time, since we only took the year 2010, and there does not appear to be a cluster effect.

For our quantitative predictors,the residuals seem to be scattered pretty evenly around the 0 line, and none of them show an obvious curving shape, so we can conclude that the linearity assumption is met for these predictors as well (Seen in figures 5,6,7,8)

For our qualitative predictors, linearity is moderately to completely satisfied (Figures  9, 10, 11 have more information).

```{r, echo = FALSE}
suicidesoutput <- augment(finalmodel) %>%
  dplyr::mutate(obs_num = row_number())
leverage_threshold <- 2*(25+1)/nrow(suicidesoutput)

leverage<- ggplot(data = suicidesoutput, aes(x = obs_num,y = .hat)) + 
geom_point(alpha = 0.7) + 
geom_hline(yintercept = leverage_threshold,color = "red")+
labs(x = "Observation Number",y = "Leverage",title = "Leverage") +
geom_text(aes(label=ifelse(.hat > leverage_threshold, as.character(obs_num), "")), nudge_x = 4)

cooks <- ggplot(data = suicidesoutput, aes(x = obs_num, y = .cooksd)) + 
geom_point(alpha = 0.7) + 
geom_hline(yintercept=1,color = "red")+
labs(x= "Observation Number",y = "Cook's Distance",title = "Cook's Distance") +
geom_text(aes(label = ifelse(.hat>1,as.character(obs_num),"")))

stand <- ggplot(data = suicidesoutput, aes(x = .fitted,y = .std.resid)) +
  geom_point(alpha = 0.7) + 
  geom_hline(yintercept = 0,color = "red") +
  geom_hline(yintercept = -2,color = "red",linetype = "dotted") +
  geom_hline(yintercept = 2,color = "red",linetype = "dotted") + labs(x ="Predicted Value",y ="Standardized Residuals",title = "Standardized Residuals vs. Predicted") + geom_text(aes(label = ifelse(abs(.std.resid) >2,as.character(obs_num),"")), nudge_x = 0.3)

plot_grid(leverage, cooks, stand)
```


There are 762 points with high leverage, meaning they affect something about the model. However, these points do not have a significant influence on the model coefficients. There are 66 observations that are considered to have standarized residuals with large magnitude, which is about 6.25% of the total number of observations. This is not a concern because this is only a small percentage greater than the standard 5%.

There are some obvious concerns with multicollinearity in this model because some of the variance inflation factors are higher than 10. Population, HDI, gdp_for_year, and gdp per capita all have variance inflation factors greater than 10. Based on the pairs plot, we can see that HDI, gdp_for_year, gdp_per_capita, and population are all very correlated. After removing HDI and GDP for year, all VIF values are below 10, meaning we do not have any multicollinearity. (Figure 12 has more information)

```{r}
suicideFinal <- suicideFinal %>% mutate (gdp_per_capita = `gdp_per_capita ($)`)

finalmodel <- lm(formula = `suicides/100k pop` ~ sex + age + population + gdp_per_capita + region + region*gdp_per_capita + age*population + age*gdp_per_capita, data = suicideFinal)
kable(tidy(finalmodel))
```

In validating our model, we see that the mean squared error values are small, so we can conclude that our model predicts well. 

## Discussion and Limitations

Our Intercept is interpreted as when the predictor variables are log(x) = 0, the expected median of suicides per 100,000 people is around  1689.14  with a baseline of female, age between 15-24 years and region of Australia and New Zealand. We will discuss this high amount in limitations.  

When population is doubled, the median of suicides/100k is expected to multiply by a factor of 0.105, which shows that increases in population correlate with lower suicide rates. In general countries a higher GDP because when when GDP / capita is doubled, the median of suicides/100k is expected to multiply by a factor of 0.93. When the gender is male, the  the median of suicides/100k is expected to multiply by a factor of 2.85 when compared to the baseline of females

We see the lowest increases in suicide rate in the Northern America model, but the greatest increases in western Europe; however, since gdp per capita interacts greatly with this those effects are limited. 

For Northern America, we also did predictions based on sex, population, and age. We empirically found that in North America, as age increases. This corroborates the general intercepts. However, since region is generally not significant, our original hypothesis regarding suicides being dependent on region is not extremely valid. 

Suicide rate in general can vary a lot depending on the year. Just looking at it by year, we noticed that some countries’ average values tend to change over time. As such, our model might not be the most accurate; however, we can predict certain things to a reasonable degree of accuracy. HDI is a limitation because you can only calculate that every 5 years. For predicting, we have to plug in numbers that make sense for a region, or else it gives us a negative grade. In terms of what we could have done better, we could have assessed multicollinearity before selecting the model, which may have changed the variables in the final model. We also could have renamed our mean-centered variables so that we wouldn’t get confused later on when trying to differentiate. If we were to conduct this project again, we would utilize a time series model, as suicide rates vary across time.

Before we checked for multicollinearity, the intercept for our final model was a reasonable number of approximately 12 suicides per 100,000, which seemed to fit the distribution of suicides per 100,000 seen in EDA. However, after removing HDI and gdp_for_year from the final model while checking for multicollinearity, the intercept shot up to approximately 1689 suicides per 100,000. This is an error that we would want to look into in the future. Additionally, the coefficient for 5-14 years is extremely high in both final models, which also does not make sense given the distribution of the number of suicides per 100,000 for each age group seen in EDA. Although the interaction effects seem to somewhat make up for this error, this is also something we would want to investigate in the future. 

We believe our regression model is appropriate given our skillset. If we could continue to work on this data set, we would see how suicide rates are changing over time and use this data to make better conclusions about future rates. 

## Conclusion

Our original hypothesis was that generally, in poorer countries we predict that suicide rates will lower. According to our final model, when GDP / capita is doubled, the median of suicides/100k is expected to multiply by a factor of .93. This proved our hypothesis to be incorrect, since when the GDP is lower, the median of suicides/100k is expected to higher than it would be when the GDP is higher. However, GDP / capita was not significant in our model, meaning it may not have a significant effect on suicides/100k. Similarly, INSERT REGION COMPARISON. However, region was found to not be significant either. We did find that population, sex, and age are all significant predictors of suicides/100k. According to our final model, suicide rates are higher for males than they are for females. Additionally, as age group increases, suicide rate increases as well. Finally, smaller populations tend to have higher suicides rates than larger populations do. Overall, we found that economic status and location do not have as large of an effect on suicide rates as population, age, and sex do.


## Additional Work

This was our original final model, but concerns of multicollienarity made us get rid of a few variables. In this original model, the intercept is much more realistic at around 7 suicides/100k given the baseline values. Concerns of multicollinearity made us get rid of two variables. However, we also believe that the interaction effects were able to counterract the high intercept and values of certain parameters. For example, we notice that for Northern America the estimate is 



###