---
title: "Stat210 Final Project"
author: "CamFam"
date: "May 1st, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-packages}
library(tidyverse)
library(dslabs)
library(nnet)
library(knitr)
library(broom)
library(skimr)
library(modelr)
library(leaps)
library(plyr)
library(dplyr)
library(rms)
```

```{r load-data}
suicide <- read_csv("/cloud/project/data/master.csv")
```

## Section 1. Introduction

With suicide being one of the leading causes of death for teens in America, we thought it would be interesting to see how the U.S. stacks up against other countries all over the world in terms of the number of suicides that occur, as well as what factors may contribute to the significant number of suicides in the U.S. and other countries.[1] We want to explore how economic status, along with variables such as age, sex, and human development index, affects suicide rates all across the world. Our hypothesis is that generally, in poorer countries we predict that suicide rates will lower. 


Our response variable will be suicides/100k pop, which is the number of suicides per 100,000 people in a certain country and year, which is stored as a numeric in our dataset. Our predictors variables will be age, sex, country, year, HDI, gdp_for_year, gdp_per_capita, generation, region and continent. Age is the age an individual was when they passed, sex is the gender of that individual, country is the country they are from, year is the year they passed, HDI for year is the human development index for a given country and year, gdp_for_year is the GDP for a given country and year, gdp_per_capita is the GDP per capita for a given country and year, and generation is the generation that an individual belongs to. We wish to understand how the number of suicides per 100,000 people in a certain country and year changes as year, GDP, GDP per capita, and HDI increase or decreases, meaning we want to understand the population coefficients for year, gdp_for_year, gdp_per_capita, and HDI for year. Additionally, we want to understand whether age, sex, generation, and country have an effect on the number of suicides per 100,000 people, meaning we also want to understand the population coefficients for these variables.

The variables relevant to the analysis of our research question are stated above: suicides/100k pop, age, sex, country, year, HDI for year, gdp_for_year, gdp_per_capita, and generation. Additionally, suicides_no (which is the number of suicides for individuals who are of a certain age group and sex and who passed in a certain country and year) and population, (which is the total number of individuals who are of a certain age group and sex and who live in a certain country in a certain year), since these are used to calculate suicides/100k pop.

We will now perform exploratory data analysis on the variables that we plan to use in our model.

Given the longitudinal structure of our data and the need for making a multilevel model, we are going to use the data collected in 2010 as that year has a majority of the countries in the origianl data set and also is the year HDI values were collected.


```{r}
suicideNew <- suicide%>% filter(year == 2010)
skim(suicideNew)
```
We first need to turn our character variables into factors.
```{r}
suicideNew <- suicideNew %>% mutate(age = as.factor(age))
suicideNew <- suicideNew %>% mutate(country = as.factor(country))
suicideNew <- suicideNew %>% mutate(generation = as.factor(generation))
suicideNew <- suicideNew %>% mutate(sex = as.factor(sex))
```

```{r}
HDINA <- suicideNew %>% filter(is.na(`HDI for year`))
```
The countries that do not have HDI calculated for them are Aruba, Puerto Rico, Korea, and Russia. Online, we can find the HDI from 2010 of South Korea and Russia, which are .884 and .780 respectively. We will do these after we add regions to our data set. 

Later on in this EDA, we will want to do a boxplot of the countries; however, that will be indiscernable based on the number of countries. Instead, we should group countries into regions and their continents. 

We will use the gapminder dataset for countries and regions. We will extract the country, continent, and region from gapMinder, change the names of differing countries, and then merge.
```{r}
gapminder2010 <- gapminder %>% filter(year == 2010)
gapminder2010 <- gapminder2010 %>% select(country, continent, region)
gapminder2010 <- data.frame(lapply(gapminder2010, as.character), stringsAsFactors=FALSE)
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="Kyrgyz Republic", "Kyrgyzstan")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="Slovak Republic", "Slovakia")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="South Korea", "Republic of Korea")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="Russia", "Russian Federation")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="St. Lucia", "Saint Lucia")
gapminder2010$country <- replace(gapminder2010$country, gapminder2010$country=="St. Vincent and the Grenadines", "Saint Vincent and Grenadines")
gapminder2010 <- data.frame(lapply(gapminder2010, as.character), stringsAsFactors=TRUE)
suicideFinal <- merge(suicideNew, gapminder2010, by = "country", all.x = TRUE)
names(suicideFinal)[names(suicideFinal) == 'HDI for year'] <- 'HDI'
```

Online, we can find the HDI from 2010 of South Korea and Russia, which are .884 and .780 respectively. 
Since the HDI's of Puerto Rico and Aruba are not publicly available, those we will find use the average HDI from the Caribbean to impute. 


```{r}
group.means<-ddply(suicideFinal,c("region"),summarise,mean=mean(HDI, na.rm = TRUE))
group.means
for(i in 1:nrow(suicideFinal)){
  if (suicideFinal$country[i] == "Republic of Korea"){
    suicideFinal$HDI[i] = .884
  }
  if (suicideFinal$country[i] == "Russian Federation"){
    suicideFinal$HDI[i] = .780
  }
  if (suicideFinal$country[i] == "Aruba" | suicideFinal$country[i] == "Puerto Rico"){
    suicideFinal$HDI[i] = .751
  }
}
```


To see the shape of the distribution of the number of suicides per 100,000 people, we can plot a histogram of the suicides/100k pop variable.

```{r}
skim(suicideFinal)
ggplot(data=suicideFinal, mapping=aes(x=`suicides/100k pop`)) + 
  geom_histogram() +
  labs(title="Distribution of the Number of Suicides per 100,000 People", x="Number of Suicides per 100,000 People")
```

```{r}
skim(suicideFinal$`suicides/100k pop`)
```

Based on this histogram, we can see that it is not normally distributed and is extremely right skewed. In fact, from skimming this variable, we can see that the mean number of suicides per 100,000 people is only 11.22, while the maximum number of suicides per 100,000 people in this dataset is 182.32. Thus, there is at least one extreme outlier in the response variable, indicating that we should perform a log transformation on the response variable. 

Since this response variable has negative infinity values (this stems from the fact that some countries have 0 suicides), we will first increase the suicide number by 1 for each country and then recalculate suicides per 100k.

```{r}
suicideFinal <- suicideFinal %>% mutate (`suicides_no` = `suicides_no`+1)
```

```{r}
suicideFinal <- suicideFinal %>% mutate (`suicides/100k pop` = `suicides_no`/(population/100000))
```

```{r}
suicideFinal <- suicideFinal %>% mutate (`suicides/100k pop` = log(`suicides/100k pop`))
```

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`suicides/100k pop`)) + 
  geom_histogram() +
  labs(title="Distribution of the Log Number of Suicides per 100,000 People", x="Log Number of Suicides per 100,000 People")
```

From log transforming suicides/100k pop, we can already see that the extreme outliers have disappeared, and the histogram seems to be approximately normally distributed. 

Note: suicides/100k pop is now log transformed for all plots below. 

We will now look at an overview of the relationships that suicides/100k pop has with each of the quantitative predictor variables (age, sex, country, year, HDI, gdp_for_year, gdp_per_capita, and generation), as well as the relationships these variables have with each other.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`age`)) + 
  geom_bar() +
  labs(title="Bar Plot of Ages", x="Age")
```
In our data set, all ages have the same number of counts. 
```{r}
ggplot(data=suicideFinal, mapping=aes(x=`sex`)) + 
  geom_bar() +
  labs(title="Bar Plot of gender", x="Sex")
```
In our data set, male and female have the same count

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`continent`)) + 
  geom_bar() +
  labs(title="Bar Plot of Continents", x="Continent")
```
In our data set, we see that most of the countries are from the Americas and Europe. 

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`region`)) + 
  geom_bar() +
  labs(title="Bar Plot of Regions", x="Region") + coord_flip()
```
From this bar graph, we notice that most of our data set seems high in regions from Western Asia, Southern Europe, South America, Northern Europe, and the Caribbean. 

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`age`, y= `suicides/100k pop`)) + 
  geom_boxplot() +
  labs(title="Age vs. Suicides", x="Age", y="Suicides")
```

By examining these boxplots, we can tell that as age increases, suicide rate tends to increase in general.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`continent`, y= `suicides/100k pop`)) + 
  geom_boxplot() +
  labs(title="Continent vs. Suicides", x="Continent", y="Suicides")
```
In general, each continent has a similar average number of suicides, with Europe having much more outliers with fewer suicides than the others. 

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`region`, y= `suicides/100k pop`)) + 
  geom_boxplot() +
  labs(title="Region vs. Suicides", x="Region", y="Suicides") + coord_flip()
```
The average number of suicides/100k for each region is around 2. We see Souther Africa as a key outlier with far less suicides and Eastern Africa with far more suicides than the average. 

```{r}
ggplot(data=suicideFinal, mapping=aes(x=sex, y=`suicides/100k pop` ))+
  geom_boxplot()+
  labs(title = "Boxplot of Gender & Suicide", x = "gender", y = "suicide")
```

From this boxplot, we notice that males have a higher suicide rate/100k people than females. However, there are many outliers in this data set so we must explore further.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=generation, y=`suicides/100k pop` ))+
  geom_boxplot()+
  labs(title = "Boxplot of Generation & Suicide", x = "Generation", y = "Suicide")
```

From this boxplot, we see that the average suicide rate/100k people is varied among generation. More specifically, we notice that Generation Z and Millienials have lower rates than the average value of around 2.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`population`)) + 
  geom_histogram() +
  labs(title="Distribution of Population", x="Population")
```
We will have to log transform population. In addition, we will mean center it so it is easier to interpret.

```{r}
suicideFinal <- suicideFinal %>% mutate(population = log(population), population = population - mean(population))
ggplot(data=suicideFinal, mapping=aes(x=`population`)) + 
  geom_histogram() +
  labs(title="Distribution of Population", x="Population")
```
We see that this plot has is skewed left and has a unimodal distribution. 

We will mean center HDI so it makes it easier to interpret. 

```{r}
suicideFinal <- suicideFinal %>% mutate(HDI = HDI - mean(HDI))
ggplot(data=suicideFinal, mapping=aes(x=`HDI`)) + 
  geom_histogram() +
  labs(title="Distribution of HDI", x="HDI")
```
The HDI does not have a set distribution. Rather, it is quite parse. 
```{r}
ggplot(data=suicideFinal, mapping=aes(x=(`gdp_for_year ($)`))) + geom_histogram() +
  labs(title="Distribution of GDPs of Countries", x="Value")
```
Because this is skewed greatly, a log transformation of gdp is needed. We will also mean center this variable. 
```{r}
suicideFinal<- suicideFinal %>% mutate (`gdp_for_year ($)` = log(`gdp_for_year ($)`))
suicideFinal <- suicideFinal %>% mutate(`gdp_for_year ($)` = `gdp_for_year ($)` - mean(`gdp_for_year ($)`))
ggplot(data=suicideFinal, mapping=aes(x=(`gdp_for_year ($)`))) + geom_histogram() +
  labs(title="Distribution of GDPs of Countries", x="Value")
```

Log transforming the data set, we notice that the distribution of GDPs of countries has a near-bimodal distribution.

```{r}
ggplot(data=suicide, mapping=aes(x=(`gdp_per_capita ($)`))) +
  geom_histogram() +
  labs(title="Distribution of GDP Per Capita of Countries", x="Value")
```
Like GDP, we will need to log transform this data set. We will also mean center this. 

```{r}
suicideFinal<- suicideFinal %>% mutate (`gdp_per_capita ($)` = log(`gdp_per_capita ($)`))
suicideFinal <- suicideFinal %>% mutate(`gdp_per_capita ($)` = `gdp_per_capita ($)` - mean(`gdp_per_capita ($)`))
ggplot(data=suicideFinal, mapping=aes(x=(`gdp_per_capita ($)`))) +
  geom_histogram() +
  labs(title="Distribution of GDP Per Capita of Countries", x="Value")
```

Like the GDP of the country, the GDP Per Capita has a non-normal distribution. Rather, it is skewed to the left and is bimodal. Since these variables are so similar, there could be mulitcollinearity between the two.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`HDI`, y=`suicides/100k pop`))+
  geom_point()+
  labs(title = "Plot of Suicides/100k vs HDI", x = "HDI", y = "Suicides/100k")
```
We do not see a correlation between HDI and Suicides/100k.
```{r}
ggplot(data=suicideFinal, mapping=aes(x=`gdp_per_capita ($)`, y=`suicides/100k pop`))+
  geom_point()+
  labs(title = "Plot of Suicides/100k vs GDP", x = "GDP", y = "Suicides/100k")
```
We do not see a correlation between GDP and Suicides/100k.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`HDI`, y=`suicides/100k pop`))+
  geom_point()+
  labs(title = "Plot of Suicides/100k vs GDP Per Capita", x = "GDP Per Capita", y = "Suicides/100k")
```
We do not see a correlation between GDP per capita and Suicides/100k.

```{r}
pairs(`suicides/100k pop` ~ `HDI` + `gdp_for_year ($)` + `gdp_per_capita ($)` + population, data = suicideFinal)
```

From the pairs plot, it looks as if  HDI, and gdp_for_year do not have a clear linear relationship with suicides/100k pop. However, gdp_per_capita seems to be positively correlated with suicides/100k pop, meaning as gdp_per_capita increases, so does suicides/100k pop. Additionally, it looks as if HDI and gdp_per_capita seem to have a strong non-linear relationship, indicating that we should continue looking into this relationship and perhaps include an interaction term between these two variables. Similarly, HDI and gdp_for_year also seem to have a strong non-linear relationship, so we should include an interaction term between these two variables as well. We also see a strong evidence of multicollinearity between GDP for year and GDP per capita that we must address in the model.

We are very concerned about multicollinearity between all of these variables, so we will look into VIF.

We plan to do a multiple linear regression because suicides/100k pop is a quantitative variable (there are no levels to it, since it is a continuous variable).  

## Section 3. Data
We are not going to include country in this model as there are over 188 countries and we can already account for much of this variability based off of region and continent. 
```{r}
model1 <- lm(`suicides/100k pop` ~ sex + age + population + HDI + (`gdp_for_year ($)`) + (`gdp_per_capita ($)`) + continent + region + generation, data = suicideFinal)
regfit_backward <- step(model1, direction = "backward")
regfit_backward
```

We chose a multiple linear regression model because we want to predict the value of a variable based on multile different independent variables. Based on backward selection, we found that sex, age  population, HDI, gdp for year, and gdp per capita, and region are relevant predictors, whereas population, continent, and generation are not. 

### Testing of Interesting Interactions

For the interaction effects, we will look into the interaction of each quantitative variable with each qualitative variable. We will also look at a select few qualitative variables with one another.

```{r}
ggplot(data=suicideFinal,mapping=aes(x=population,y=`suicides/100k pop`,color=as.factor(region))) +
geom_smooth(method="lm", se=FALSE) + labs(title="Relationship between suicides/100k pop and population with regards to region", 
x="population", 
y="suicides",
color="region")

ggplot(data=suicideFinal,mapping=aes(x=HDI,y=`suicides/100k pop`,color=as.factor(region))) +
geom_smooth(method="lm", se=FALSE) + labs(title="Relationship between suicides/100k pop and HDI with regards to region", 
x="HDI", 
y="suicides",
color="region")

ggplot(data=suicideFinal,mapping=aes(x=`gdp_per_capita ($)`,y=`suicides/100k pop`,color=as.factor(region))) +
geom_smooth(method="lm", se=FALSE) + labs(title="Relationship between suicides/100k pop and GDP / capita with regards to region view", 
x="GDP / capita", 
y="suicides",
color="region")

ggplot(data=suicideFinal,mapping=aes(x=`gdp_for_year ($)`,y=`suicides/100k pop`,color=as.factor(region))) +
geom_smooth(method="lm", se=FALSE) + labs(title="Relationship between suicides/100k pop and GDP / year with regards to region view", 
x="GDP / year", 
y="suicides",
color="region")
```

There seems to be a significant interaction effect between all of the quantitative variables that were selected in backwards selection (population, GDP per capita, GDP for year, and HDI) and region because the lines are not relatively parallel, as the slopes of the all of the lines in each graph look to be quite different. Thus, we will use Nested F Tests to formally test if any of these interactions are significant.

```{r}
model2 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + region*HDI, data = suicideFinal)
anova(model1,model2)
```
The F statistic is 2.4591 and the p-value is 0.003639, which is  small. We can conclude that the interaction effect of HDI and region is a significant predictor.

```{r}
model3 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + region*population, data = suicideFinal)
anova(model1,model3)
```
The F statistic is 4.3923 and the p-value is 4.641e-08, which is very low. We can conclude that the interaction effect of region and population is  a significant predictor.

```{r}
model4 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + region*`gdp_for_year ($)`, data = suicideFinal)
anova(model1,model4)
```
The F statistic is 2.216 and the p-value is 0.007552 , which is below .05. We can conclude that the interaction effect of region and GDP for year is  a significant predictor.

```{r}
model5 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + region*`gdp_per_capita ($)`, data = suicideFinal)
anova(model1,model5)
```
The F statistic is 1.2491 and the p-value is 0.2387, which is above .05. Contrary to what we first thought, we can conclude that the interaction effect of region and gdp per capita is not a significant predictor.

```{r}
model6 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + sex*HDI, data = suicideFinal)
anova(model1,model6)
```
We see that sex will not have a significant effect on any interactions. We will not be looking at the interaction effects between sex and other variables.

```{r}
model10 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + age*HDI, data = suicideFinal)
anova(model1,model10)
```
The F statistic is 13.857 and the p-value is 4.027e-13, which is smaller than 0.05. We can conclude that the interaction effect of age and HDI is significant predictor.

```{r}
model11 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + age*`gdp_per_capita ($)`, data = suicideFinal)
anova(model1,model11)
```
The F statistic is 3.0279 and the p-value is 0.01015, which is smaller than 0.05. We can conclude that the interaction effect of age and gdp per capita is a significant predictor.

```{r}
model12 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + age*`gdp_for_year ($)`, data = suicideFinal)
anova(model1,model12)
```
The F statistic is 27.466 and the p-value is 2.2e-16, which is smaller than 0.05. We can conclude that the interaction effect of age and gdp for year is significant predictor.

```{r}
model13 <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + age*population, data = suicideFinal)
anova(model1,model13)
```
The F statistic is 23.793 and the p-value is 2.2e-16, which is smaller than 0.05. We can conclude that the interaction effect of age and gdp per capita is a significant predictor.

From this, analysis, we will include 7 interaction effects in our new model
```{r}
finalmodel <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + region*HDI + region*`gdp_for_year ($)` + region*`gdp_per_capita ($)` + age*HDI + age*population + age*`gdp_per_capita ($)` + age*`gdp_for_year ($)`, data = suicideFinal)
kable(tidy(finalmodel))
```

```{r}
finalmodel <- lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region + region*HDI + region*`gdp_for_year ($)` + region*`gdp_per_capita ($)` + age*HDI + age*population + age*`gdp_per_capita ($)` + age*`gdp_for_year ($)`, data = suicideFinal)
kable(tidy(finalmodel))
```

### Model Assumptions

```{r}
suicideFinal <- suicideFinal %>%
  mutate(predicted = predict.lm(finalmodel), resid = residuals(finalmodel))
```

```{r}
ggplot(data=suicideFinal, mapping=aes(x=predicted,y=resid)) +
  geom_point() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. Predicted Values",
       x="Predicted Values", 
       y="Residuals")
```

From the residual vs. predicted values, the residuals with lower predicted values seem to be more sparse and spread out than the others, which may be worth looking into. However, for the most part, the majority of the residuals are randomly distributed around the red line and do not exhibit any obvious nonlinear trends. Thus, we can conclude that constant variance is satisfied.

We will now look at predictors vs. residuals to more closely examine linearity.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=region,y=resid)) +
  geom_boxplot() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. Regions",
       x="Regions", 
       y="Residuals") + coord_flip()
```

While there are some differences between some of the means of the residuals for some regions, these means do not seem to vary by too much, and thus the linearity assumption is moderately satisfied for this variable. However, there are a few regions with some extreme residuals, such as the Carribean and Western Asia. It could be worth looking into these more.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=sex,y=resid)) +
  geom_boxplot() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. Sex",
       x="Sex", 
       y="Residuals") + coord_flip()
```

While there is some differences between the means of the residuals for the two gender groups, these means do not seem to vary by too much, and the plots seem to have very similar distributions, and thus the linearity assumption is moderately satisfied for this variable.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=age,y=resid)) +
  geom_boxplot() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. Age",
       x="Age", 
       y="Residuals") + coord_flip()
```

Again, while there are some differences between some of the means of the residuals for some age groups, these means do not seem to vary by too much, and thus the linearity assumption is satisfied for this variable.

```{r}
ggplot(data=suicideFinal, mapping=aes(x=population,y=resid)) +
  geom_point() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. Population (Centered)",
       x="Population (Centered)", 
       y="Residuals")
```

```{r}
ggplot(data=suicideFinal, mapping=aes(x=HDI,y=resid)) +
  geom_point() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. HDI (Centered)",
       x="HDI (Centered)", 
       y="Residuals")
```

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`gdp_for_year ($)`,y=resid)) +
  geom_point() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. GDP for Year (Centered)",
       x="GDP for Year (Centered)", 
       y="Residuals")
```

```{r}
ggplot(data=suicideFinal, mapping=aes(x=`gdp_per_capita ($)`,y=resid)) +
  geom_point() + 
  geom_hline(yintercept=0,color="red")+
  labs(title="Residuals vs. GDP per Capita (Centered)",
       x="GDP per Capita (Centered)", 
       y="Residuals")
```

For all of these scatterplots, the residuals seem to be scattered pretty evenly around the 0 line, and none of them show an obvious curving shape, so we can conclude that the linearity assumption is met for these predictors as well.

We also need to check the normality and independence assumptions. For normality, we can create a QQ-plot of the residuals and a histogram of the distribution of residuals for this model.

```{r}
ggplot(data=suicideFinal, mapping=aes(sample=resid)) +
  stat_qq() + 
  stat_qq_line()+
labs(title="Normal QQ Plot of Residuals for the Suicide Rate Dataset")
```

```{r message = FALSE}
ggplot(data=suicideFinal, mapping=aes(x=resid)) +
  geom_histogram() +  
  labs(title="Distribution of the Residuals for the Suicide Rate Dataset", x="Residuals")
```

According to the QQ-plot, the normality assumption seems to be satisfied because the majority of the points seem to align well with the expected QQ-plot line. In addition, the histogram seems to be approximately normal and follows a mostly smooth curve.

Finally, based on the description of the data, the independence assumption seems to be met because this data was not collected over time, since we only took the year 2010, and there does not appear to be a cluster effect.

### Leverage

```{r}
suicidesoutput <- augment(finalmodel) %>%
  dplyr::mutate(obs_num = row_number())

leverage_threshold <- 2*(25+1)/nrow(suicidesoutput)

ggplot(data = suicidesoutput, aes(x = obs_num,y = .hat)) + 
geom_point(alpha = 0.7) + 
geom_hline(yintercept = leverage_threshold,color = "red")+
labs(x = "Observation Number",y = "Leverage",title = "Leverage") +
geom_text(aes(label=ifelse(.hat > leverage_threshold, as.character(obs_num), "")), nudge_x = 4)
```


```{r}
suicidesoutput %>% filter(.hat > leverage_threshold) %>% 
  select(region, population, age, sex, HDI, gdp_for_year...., gdp_per_capita....)
```

There are 762 points with high leverage.

```{r}
ggplot(data = suicidesoutput, aes(x = obs_num, y = .cooksd)) + 
geom_point(alpha = 0.7) + 
geom_hline(yintercept=1,color = "red")+
labs(x= "Observation Number",y = "Cook's Distance",title = "Cook's Distance") +
geom_text(aes(label = ifelse(.hat>1,as.character(obs_num),"")))

```

Based on cook’s distance, these points do not have a significant influence on the model coefficients. None of these points have a cook’s distance that is greater than the threshold of 1.

```{r}
ggplot(data = suicidesoutput, aes(x = .fitted,y = .std.resid)) +
  geom_point(alpha = 0.7) + 
  geom_hline(yintercept = 0,color = "red") +
  geom_hline(yintercept = -2,color = "red",linetype = "dotted") +
  geom_hline(yintercept = 2,color = "red",linetype = "dotted") + labs(x ="Predicted Value",y ="Standardized Residuals",title = "Standardized Residuals vs. Predicted") + geom_text(aes(label = ifelse(abs(.std.resid) >2,as.character(obs_num),"")), nudge_x = 0.3)
```

```{r}
suicidesoutput %>% filter(abs(.std.resid) > 2) %>% 
  select(region, population, age, sex, HDI, gdp_for_year...., gdp_per_capita....)
```

There are 66 observations that are considered to have standarized residuals with large magnitude.

Based on a N(0,1) distribution, we expect 5% of observations to have standardized residuals with magnitude greater than 2 because of the 68-95-99.7 rule.

There are 66 observations flagged as having standardized residuals with large magnitude and 1056 observations in total, so approximately 6.25% of observations have standardized residuals with magnitude greater than 2. This is not a concern because this is only a small percentage greater than 5%.

```{r}
tidy(vif(finalmodel))
```

Because all of the values are NA, this means these variables are perfectly or almost perfectly correlated. This is most likely to our inclusion of interaction terms in our model. Thus, we will remove these terms and run VIF again. 

```{r}
finalmodel1 <-lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_for_year ($)` + `gdp_per_capita ($)` + region , data = suicideFinal)
tidy(vif(finalmodel1))
```

There are some obvious concerns with multicollinearity in this model because the threshold for the variance inflation factor is 10, and some of the variance inflation factors are higher than 10. Population, HDI, `gdp_for_year ($)`, and `gdp_for_year ($)` all have variance inflation factors greater than 10.

```{r}
pairs(`suicides/100k pop` ~ `HDI` + `gdp_for_year ($)` + `gdp_per_capita ($)` + population, data = suicideFinal)
```

Based on this pairs plot from EDA, we can see that HDI, gdp_for_year, gdp_per_capita, and population are all very correlated. Thus, we will remove gdp_for_year from the model, since it seems to have the strongest correlation with all three of the other variables, and run VIF again. 

```{r}
finalmodel2 <-lm(formula = `suicides/100k pop` ~ sex + age + population + HDI + `gdp_per_capita ($)` + region , data = suicideFinal)
tidy(vif(finalmodel2))
```

All predictors now have variance inflation factors less than 10 except for HDI. Thus, we can remove HDI as well from the model.

```{r}
finalmodel3 <-lm(formula = `suicides/100k pop` ~ sex + age + population + `gdp_per_capita ($)` + region , data = suicideFinal)
tidy(vif(finalmodel3))
```

Now we see that all VIF values are below 10, meaning that we do not have any multicollinearity. Adding back the interaction terms that still have both main effects in the model, our final model is: 

```{r}
finalmodel <- lm(formula = `suicides/100k pop` ~ sex + age + population + `gdp_per_capita ($)` + region + region*`gdp_per_capita ($)` + age*population + age*`gdp_per_capita ($)`, data = suicideFinal)
kable(tidy(finalmodel))
```

### Model Validation 

We are again removing our interaction terms in this part because the map() function cannot calculate 'models' with the interaction terms in the model.

```{r}
set.seed(04012019)
mydata_cv <- crossv_kfold(suicideFinal, 5)
models <- map(mydata_cv$train, ~ lm(`suicides/100k pop` ~ sex + age + population + `gdp_per_capita ($)` + region, data = .))
test_mse <- map2_dbl(models, mydata_cv$test, mse)
test_mse
```

All of these values are relatively small, so we can conclude that our model predicts well.

### Section 3: Discussion and Limitations

### Section 4 - References
- [1] https://www.cdc.gov/nchs/fastats/adolescent-health.htm
- [2] http://hdr.undp.org/en/content/human-development-index-hdi
Your project goes here! Before you submit, make sure your chunks are turned off with `echo = FALSE`. 

You can add sections as you see fit. At a minimum, you should have the following sections: 

- Section 1: Introduction (includes introduction and exploratory data analysis)
- Section 2: Regression Analysis (includes the final model and discussion of assumptions)
- Section 3: Discussion and Limitations 
- Section 4: Conclusion
- Section 5: Additional Work