---
title: "The Effect of Socio-Economics on Suicide Rates"
author: "Cam Fam"
date: "March 29, 2019"
output: github_document
---

```{r load-packages}
library(nnet)
library(tidyverse)
library(knitr)
library(broom)
```


## Section 1. Introduction

With suicide being one of the leading causes of death for teens in America, we thought it would be interesting to see how the U.S. stacks up against other countries all over the world in terms of the number of suicides that occur, as well as what factors may contribute to the significant number of suicides in the U.S. and other countries.[1] We want to explore how economic status, along with variables such as age, sex, and human development index, affects suicide rates all across the world. Our hypothesis is that generally, in poorer countries we predict that suicide rates will lower. 

## Section 2. Analysis plan

Our response variable will be suicides/100k pop, which is the number of suicides per 100,000 people in a certain country and year, which is stored as a numeric in our dataset. Our predictors variables will be age, sex, country, year, HDI for year, gdp_for_year, gdp_per_capita, and generation. Age is the age an individual was when they passed, sex is the gender of that individual, country is the country they are from, year is the year they passed, HDI for year is the human development index for a given country and year, gdp_for_year is the GDP for a given country and year, gdp_per_capita is the GDP per capita for a given country and year, and generation is the generation that an individual belongs to. We wish to understand how the number of suicides per 100,000 people in a certain country and year changes as year, GDP, GDP per capita, and HDI increase or decreases, meaning we want to understand the population coefficients for year, gdp_for_year, gdp_per_capita, and HDI for year. Additionally, we want to understand whether age, sex, generation, and country have an effect on the number of suicides per 100,000 people, meaning we also want to understand the population coefficients for these variables.

The variables relevant to the analysis of our research question are stated above: suicides/100k pop, age, sex, country, year, HDI for year, gdp_for_year, gdp_per_capita, and generation. Additionally, suicides_no (which is the number of suicides for individuals who are of a certain age group and sex and who passed in a certain country and year) and population, (which is the total number of individuals who are of a certain age group and sex and who live in a certain country in a certain year), since these are used to calculate suicides/100k pop.

EDA.

We plan to do a multiple linear regression because suicides/100k pop is a quantitative variable (there are no levels to it, since it is a continuous variable). 

Based on the variables, we plan to predict suicide rate in the future for certain countries, as well as what generation someone who commits suicide is in given other variables.

## Section 3. Data

```{r load-data}
suicide <- read_csv("/cloud/project/data/master.csv")
```

```{r}
glimpse(suicide)
```

### Section 4 - References
- [1] https://www.cdc.gov/nchs/fastats/adolescent-health.htm
- [2] http://hdr.undp.org/en/content/human-development-index-hdi
